---
title: Stock Comparison
author: "Your Name"
date: "Today's Date"
output: 
  html_document:
    toc: true
    toc_float: true
params:
  symbol1: "FB"
  symbol2: "GOOG"
---

# Introduction

```{r include = FALSE}
library(quantmod)
library(dygraphs)

stock1 <- getSymbols(params$symbol1, auto.assign = FALSE,
                    from = "2016-01-01", end = Sys.Date())
stock2 <- getSymbols(params$symbol2, auto.assign = FALSE,
                    from = "2016-01-01", end = Sys.Date())
```


This report examines the price of two stocks: 
  
* `r params$symbol1`
* `r params$symbol2` 

It is written in [R Markdown](http://rmarkdown.rstudio.com), an authoring format that enables easy creation of dynamic documents, presentations, and reports from R. R Markdown combines the core syntax of **markdown** (an easy to write plain text format) with embedded **R** code chunks that are run so their output can be included in the final document.

The data comes from [Yahoo finance](http://finance.yahoo.com).

### `r params$symbol1` price

```{r echo = FALSE}
dygraph(Cl(stock1))
```

### `r params$symbol2` price

```{r echo = FALSE}
dygraph(Cl(stock2))
```

### `r params$symbol1` volume

```{r echo = FALSE}
dygraph(Vo(stock1))
```

### `r params$symbol2` volume

```{r echo = FALSE}
dygraph(Vo(stock2))
```

