---
title: "Verbatim Chunks in R Markdown"
author: "Ramnath Vaidyanathan"
date: "March 12, 2014"
output:
  html_document:
    toc_float: false
---

_This article is reposted from [Addicted to R](https://ramnathv.github.io/posts/verbatim-chunks-knitr/) with the kind permission of Ramnath Vaidyanathan._

```{r echo = F, cache = F}
knitr::opts_chunk$set(tidy = F)
```


[RMarkdown](https://www.rstudio.com/ide/docs/r_markdown) has truly revolutinized the way I write. The simplicity of [markdown](http://daringfireball.net/projects/markdown/) and the extreme flexibility of [knitr](http://github.com/yihui/knitr) allows me to combine text and code in ways I would never have been able to do before.

One of the things I have always wanted to do is to be able to print code chunks verbatim, along with the chunk markup. This is especially useful while writing tutorials, where I want the user to see the code chunk "asis". 

Initially, I thought that I could achieve this by indenting the code chunk, thereby "fooling" [knitr](http://github.com/yihui/knitr) into believing that it is not a code chunk, and thereby printing it verbatim in a `<code>` block. But I was terribly mistaken. Apparently, `knitr` is too smart to be fooled, and no matter how deep you tuck away or indent the chunk, `knitr` always finds it and executes it!

Digging a little deeper, I found an example from the author in his [knitr-examples](https://github.com/yihui/knitr-examples/blob/master/062-chunk-wrapper.Rmd) repo. The basic idea was to write a [chunk hook](yihui.name/knitr/hooks) that wraps the source code with the markup. It makes use of the option `params.src` that contains the chunk options that are set locally.

```{r setup}
knitr::knit_hooks$set(wrapper = function(before, options, envir) {
  if (before) {
    sprintf('    ```{r %s}\n', options$params.src)
  } else '    ```\n'
})
```

Let us test it out by adding `wrapper = TRUE` to a code chunk.

```{r test_label, wrapper=TRUE, eval=FALSE}
1+1
plot(1)
```

It works! Well, it almost works. I say almost, because I don't like how it splits the code and the chunk markup into different blocks. I am sure that I can clean this up using some regex-fu, but I wanted to explore a cleaner solution using hooks.

After a lot of experimentation, I ended up with the following hook. Note that my implementation involves a source hook and not a chunk hook, like the one proposed by [Yihui](http://yihui.name). It is customized for `Rmd` documents, but is easy to extend for other formats like `Rnw`.

```{r cache = F}
knitr::knit_hooks$set(source = function(x, options){
  if (!is.null(options$verbatim) && options$verbatim){
    opts = gsub(",\\s*verbatim\\s*=\\s*TRUE\\s*", "", options$params.src)
    bef = sprintf('\n\n    ```{r %s}\n', opts, "\n")
    stringr::str_c(
      bef, 
      knitr:::indent_block(paste(x, collapse = '\n'), "    "), 
      "\n    ```\n"
    )
  } else {
    stringr::str_c("\n\n```", tolower(options$engine), "\n", 
      paste(x, collapse = '\n'), "\n```\n\n"
    )
  }
})
```

Let us now test how this hook performs.

```{r qplot, fig.width = 5, fig.height = 5, fig.align = "center", verbatim = TRUE, message=FALSE}
library(ggplot2)
qplot(wt, am, color = vs, data = mtcars)
```

Yeyy! It works exactly like how I wanted it to. As always, this post is fully reproducible, and you can find the code [here](https://ramnathv.github.io/posts/verbatim-chunks-knitr/index.Rmd).


<style>
pre {
  padding: 0;
}
pre code {
  padding: 1rem;
}
p {
  text-align: justify;
}
</style>
