---
title: "HTML Widgets"
---

HTML widgets enable you to add new types of HTML output to R Markdown documents. You might want to do this to add a binding between R and a Javascript data visualization library like [d3](http://d3js.org) or [Leaflet](http://leafletjs.com). HTML widgets create output that works both inside R Markdown documents as well as while executing code from the console (e.g. to display it the RStudio Viewer).

#### Prerequisties

Creating HTML widgets requires the development versions of both the **knitr** and **htmltools** packages. You can install them as follows:

```{r, eval=FALSE}
devtools::install_github(c("yihui/knitr", "rstudio/htmltools"))
```

## Creating an HTML Widget

To create an HTML widget you define an R function that does the following:

1) Dynamically generates HTML based on it's input parameters;

2) Enumerates the dependencies (e.g. CSS and Javascript) required by that HTML; and

3) Returns the HTML and dependencies as an object of class "html_output"

When implementing an HTML widget you use the `html_dependency` and `html_output` functions from the **htmltools** package. The basic skeleton of a function that defines a new widget looks like this:


```{r, eval=FALSE}
snazzy_visualization <- function(data, width, height) {
  
  # generate html for the visualization
  html <- generate_snazzy_html(data, width, height)
  
  # define dependencies
  dependencies <- list(
    htmltools::html_dependency(...),
    htmltools::html_dependency(...)
  )
  
  # return output
  htmltools::html_output(html, dependencies)  
}
```

Dependencies are defined with a name, version, path and optional meta, script, and stylesheet elements to include within the document head (see below for examples). Multiple instances of a dependency with the same name (e.g. "jquery") are merged into a single dependency and the most recent version specified is used.

It's important to note that dependencies must be local rather than remote (i.e. they must exist on the local filesystem). This requirement exists so that HTML documents created with R Markdown can be self contained and always work offline.

## Examples

Here are two examples of HTML widgets. These examples are from the **rmdexamples** package located at <https://github.com/rstudio/rmdexamples>.

### JustGage

[JustGage](http://justgage.com/) is Javascript library used to create numeric gauge visualizations. JustGage depends on both it's on Javascript as well as the [raphael.js](http://raphaeljs.com/) library. Here's what a function that defines a JustGage HTML widget might look like:

```{r, eval=FALSE}
justgage <- function(title, value, min, max, 
                     label = NULL, width = 200, height = 160) {
  
  # create random/unique id to bind the div and script
  id <- paste("justgage", as.integer(stats::runif(1, 1, 10000)), sep="-") 

  # create a style attribute for the width and height
  style <- paste("width:", width, "px;height:", height, "px", sep = "")
  
  # create a list representing the parameters to JustGage
  options <- list(id = id,
                  title = title,
                  value = value,
                  min = min,
                  max = max,
                  label = label)

  # generate html for the justgage
  html <- paste(
    "<div id=\"", id, "\" ",  "style=\"", style, "\">", "</div>", 
    "<script>var g = new JustGage(", RJSONIO::toJSON(options), ");</script>",
    sep = "")
  
  # define html dependencies
  dependencies <- list(
    htmltools::html_dependency(
      name = "raphael",
      version = "2.1.2",
      path = system.file("www/libs/raphael-2.1.2", package = "rmdexamples"),
      script = "raphael.js"
    ),
    htmltools::html_dependency(
      name = "justgage",
      version = "1.0.1",
      path = system.file("www/libs/justgage-1.0.1", package = "rmdexamples"),
      script = "justgage.1.0.1.min.js"
    )
  )
  
  # return html output
  htmltools::html_output(html, dependencies)
}
```

Note that the required Javascript libraries are contained within the function's package ([rmdexamples](https://github.com/rstudio/rmdexamples)) so are accessed using `system.file`.

### DataTables

[DataTables](https://datatables.net/) is a plug-in for the [jQuery](http://jquery.com/) Javascript library that adds advanced interaction controls to any HTML table. Here's what a function that defines a DataTables widget might look like:

```{r, eval=FALSE}
datatable <- function(x, digits = getOption("digits"), 
                      row.names = NA, align = NULL, ...) {
  
  # create random/unique id for the table
  id <- paste("datatable", as.integer(stats::runif(1, 1, 10000)), sep="-") 
  
  # generate an html version of the table that includes the id
  html <- knitr::kable(
    x, 
    format = "html", 
    digits = digits,
    row.names = NA,
    align = align,
    output = FALSE,
    table.attr = paste("id=\"", id, "\" ", 
                        "class=\"table table-bordered\" ",
                        sep = ""),
    ...)
  
  # create the script which binds the datatable
  html <- paste(html,
    "<script>",
      "$(document).ready(function() {",
        "$('#", id ,"').dataTable();",
      "});",
   "</script>", sep = "")
  
  # define dependencies
  dependencies <- list(
    htmltools::html_dependency(
      name = "jquery",
      version = "1.11.0",
      path = system.file("www/libs/jquery-1.11.0", package = "rmdexamples"),
      script = "jquery.min.js"
    ),
    htmltools::html_dependency(
      name = "datatables",
      version = "1.9.4",
      path = system.file("www/libs/datatables-1.9.4", package = "rmdexamples"),
      stylesheet = "css/jquery.dataTables.css",
      script = "js/jquery.dataTables.min.js"
    )
  )
  
  # return html output
  htmltools::html_output(html, dependencies)
}
```

Note that the `knitr::kable` function is used to generate the HTML table and that the paramters to the `datatable` function are passed through to it. Note also that the required Javascript libraries are contained within the function's package ([rmdexamples](https://github.com/rstudio/rmdexamples)) so are accessed using `system.file`.




