---
title: "HTML Widgets"
output: html_document
---

HTML widgets enable you to add new types of HTML output to R Markdown documents. You might want to do this to add a binding between R and a Javascript data visualization library like [d3](http://d3js.org) or [Leaflet](http://leafletjs.com). HTML widgets create output that works both inside R Markdown documents as well as while executing code from the console (e.g. to display within the RStudio Viewer).

#### Prerequisties

Creating HTML widgets requires the development versions of both the **knitr** and **htmltools** packages. You can install them as follows:

```{r, eval=FALSE}
devtools::install_github(c("yihui/knitr", "rstudio/htmltools"))
```

## Creating an HTML Widget

HTML widgets are "printed" into various contexts using S3 method dispatch. A typical widget supports printing in two different contexts:

1) At the R console (where printing occurs to the RStudio Viewer pane or within an external web browser)

2) Within R Markdown documents (where printing occurs inline within the document)

To accomplish this, HTML widgets implement the `print` and `knit_print` S3 methods. Printing in both contexts includes not only the raw HTML but also the specification of dependencies (e.g. Javascript libraires).

The **htmltools** package includes helper functions that make it straightforward to implement both of these print methods and share the underlying code used to generate HTML and specify dependencies (the two examples below illustrate the use of these functions).

### Generating HTML

A couple of important things to keep in mind when generating HTML for use in widgets:

1) Due to the way R Markdown processes text within HTML tags you should not use indendation in your generated HTML. This is because pandoc interprets any content indented 4 spaces as a code / preformatted block.

2) Text included within attributes or the body of tags should be escaped. You can use the `htmltools::html_escape` function to do this (the DataTables example below illustrates this).

3) Literals within JSON also needs to be escaped. The best way to do this (and to ensure your JSON is fully valid) is to create an R list and then use `RJSONIO::toJSON` to generate the JSON (the JustGage example below illustrates this).

### Specifying Dependencies

HTML dependencies are defined using the `htmltools::html_dependency` function and include a name, version, path and optional meta, script, and stylesheet elements to include within the document head. For example:

```{r, eval=FALSE}
htmltools::html_dependency(
  name = "jquery",
  version = "1.11.0",
  path = system.file("www/libs/jquery-1.11.0", package = "rmdexamples"),
  script = "jquery.min.js"
)
```

Multiple instances of a dependency with the same name (e.g. "jquery") are merged into a single dependency and the most recent version specified is used.

It's important to note that dependencies must be local rather than remote (i.e. they must exist on the local filesystem). This requirement exists so that HTML documents created with R Markdown can be self contained and always work offline.

## Examples

Here are two examples of HTML widgets. These examples are from the **rmdexamples** package located at <https://github.com/rstudio/rmdexamples>.

### JustGage

[JustGage](http://justgage.com/) is Javascript library used to create numeric gauge visualizations. JustGage depends on both it's on Javascript as well as the [raphael.js](http://raphaeljs.com/) library. Here's what an HTML widget for JustGage might look like:

```{r, eval=FALSE}
#' @export
justgage <- function(title, value, min, max, label = NULL) {
  structure(class = "justgage", list(
    title = title,
    label = label,
    value = value,
    min = min,
    max = max
  ))
}

#' @export
print.justgage <- function(x, ...) {
  htmltools::html_print(
    justgage_html(x, 450, 350), 
    justgage_dependencies()
  )
}

#' @export
knit_print.justgage <- function(x, options) {
  htmltools::html_knit_print(
    justgage_html(x, options$out.width.px, options$out.height.px), 
    justgage_dependencies()
  )
}

justgage_html <- function(x, width, height) {
  
  # create random/unique id to bind the div and script
  id <- paste("justgage", as.integer(stats::runif(1, 1, 10000)), sep="-") 
  
  # create a style attribute for the width and height
  style <- paste("width:", width, "px;height:", height, "px", sep = "")
  
  # create a list representing the parameters to JustGage
  options <- list(id = id,
                  title = x$title,
                  value = x$value,
                  min = x$min,
                  max = x$max,
                  label = x$label)

  # generate html for the justgage
  html <- paste(
    "<div id=\"", id, "\" ",  "style=\"", style, "\">", "</div>", 
    "<script>var g = new JustGage(", RJSONIO::toJSON(options), ");</script>",
    sep = "")
  
  # return html
  html
}

justgage_dependencies <- function() {
  list(
    htmltools::html_dependency(
      name = "raphael",
      version = "2.1.2",
      path = system.file("www/libs/raphael-2.1.2", package = "rmdexamples"),
      script = "raphael.js"
    ),
    htmltools::html_dependency(
      name = "justgage",
      version = "1.0.1",
      path = system.file("www/libs/justgage-1.0.1", package = "rmdexamples"),
      script = "justgage.1.0.1.min.js"
    )
  )
}
```

Note that knitr chunk options are consulted to determine the size of the JustGage to output. Note also that the required Javascript libraries are contained within the function's package ([rmdexamples](https://github.com/rstudio/rmdexamples)) so are accessed using `system.file`.

### DataTables

[DataTables](https://datatables.net/) is a plug-in for the [jQuery](http://jquery.com/) Javascript library that adds advanced interaction controls to any HTML table. Here's what an HTNL widget for DataTables might look like:

```{r, eval=FALSE}
#' @export
datatable <- function(data, align = NULL) {
  structure(class = "datatable", list(
    data = data,
    align = align
  ))
}

#' @export
print.datatable <- function(x, ...) {
  htmltools::html_print(datatable_html(x), datatable_dependencies())
}

#' @export
knit_print.datatable <- function(x, options) {
  htmltools::html_knit_print(datatable_html(x), datatable_dependencies())
}

datatable_html <- function(x) {
  
  # create random/unique id for the table
  id <- paste("datatable", as.integer(stats::runif(1, 1, 10000)), sep="-") 
  
  # generate an html version of the table that includes the id
  data <- x$data
  align = if (is.null(align <- x$align)) '' else {
    sprintf(' align="%s"', c(l = 'left', c = 'center', r = 'right')[align])
  }
  html <- paste(c(
    sprintf("<table id = \"%s\">", id),
    c('<thead>', '<tr>', 
      sprintf('<th>%s</th>', htmltools::html_escape(colnames(data))), 
      '</tr>', '</thead>'),
    '<tbody>',
    paste(
      '<tr>',
      apply(data, 1, function(z) {
        z = htmltools::html_escape(z)
        paste(sprintf('<td%s>%s</td>', align, z), collapse = '')
      }),
      '</tr>', sep = ''
    ),
    '</tbody>',
    '</table>'
  ), sep = '', collapse = '')
   
  # append the script which binds the datatable
  html <- paste(html,
                "<div>&nbsp;</div>",
                "<script>",
                "$(document).ready(function() {",
                "$('#", id ,"').dataTable();",
                "});",
                "</script>", sep = "")
    
  # return html
  html
}

datatable_dependencies <- function() {
  list(
    htmltools::html_dependency(
      name = "jquery",
      version = "1.11.0",
      path = system.file("www/libs/jquery-1.11.0", package = "rmdexamples"),
      script = "jquery.min.js"
    ),
    htmltools::html_dependency(
      name = "datatables",
      version = "1.9.4",
      path = system.file("www/libs/datatables-1.9.4", package = "rmdexamples"),
      stylesheet = "css/jquery.dataTables.css",
      script = "js/jquery.dataTables.min.js"
    )
  )
}
```

Note that the required Javascript libraries are contained within the function's package ([rmdexamples](https://github.com/rstudio/rmdexamples)) so are accessed using `system.file`.





