---
title: "Code Languages"
output:
  html_document:
    toc_float: false
    includes:
      before_body: [include_header.html, include_lessons_nav.html]
      after_body: [include_lessons_nav_end.html, include_footer.html]
---

In addition to executing R code chunks, the [knitr](http://yihui.name/knitr/) package can also execute chunks in a variety of other languages. Some of the available language engines include:

* Python
* SQL
* Bash
* Rcpp
* Stan
* JavaScript
* CSS

To process a code chunk using an alternate language engine you simply use the name of the engine in place of `r` in your chunk declaration, for example:

```{r echo = FALSE, comment = ""}
cat(htmltools::includeText("demos/bash.Rmd"))
```

## Example

The python engine enables execution of python code via an external python interpreter (the interpreter returned by `Sys.which("python")` is used to execute the code). Here’s a simple example:

```{r echo = FALSE, comment = ""}
cat(htmltools::includeText("demos/python.Rmd"))
```

```{python, echo = FALSE}
x = 'hello, python world!'
print(x.split(' '))
```

Note that chunk options like `echo` and `results` are all valid when using a language engine like python. 

Learn more about using other languages with R Markdown in [knitr Language Engines](authoring_knitr_engines.html).

***

## [Continue to Parameters](lesson-6.html)
