---
title: "Code Chunks"
output:
  html_document:
    toc_float: false
    includes:
      before_body: [include_header.html, include_lessons_nav.html]
      after_body: [include_lessons_nav_end.html, include_footer.html]
---

To add an R code chunk to your .Rmd file, type your code between the lines

```{r echo = FALSE, comment = ""}
cat(htmltools::includeText("demos/chunk-boundaries.Rmd"))
```

When you render your .Rmd file, R Markdown will run each code chunk and embed the results beneath the code chunk in your final report.

<!-- Screenshot of 2-chunks.Rmd being rendered.
![]()
-->

## Chunk Options

Add [knitr options](http://yihui.name/knitr/options/) to a chunk header to customize how the contents are displayed.

In the [file above](demos/2-chunks.Rmd), the option `echo=FALSE` prevents a chunk's code from appearing in the final report. `results='hide'` prevents a chunk's results from appearing in the final report.

You can view a quick list of knitr chunk options in the [R Markdown Reference Guide](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf).

## Global Options

To set global options that apply to every chunk in your file, call `knitr::opts_chunk$set` from a code chunk, e.g

```{r echo = FALSE, comment = ""}
cat(htmltools::includeText("demos/global-chunk-options.Rmd"))
```

Knitr will treat each option that you pass to `knitr::opts_chunk$set` as a global default that can be overwritten in individual chunk headers.

## Table Output

By default, data frames and matrixes are output as they would be in the R terminal (in a monospaced font). If you prefer that data be displayed with additional formatting you can use the `knitr::kable` function. For example:

```{r echo = FALSE, comment = ""}
cat(htmltools::includeText("demos/kable.Rmd"))
```

```{r, echo = FALSE, results='asis'}
knitr::kable(mtcars[1:5, ])
```

Note the use of the `results='asis'` chunk option. This is required to ensure that the raw table output isn’t processed further by knitr.

## Caching

If document rendering becomes time consuming due to long computations you can use knitr caching to improve performance. The documentation [knitr chunk and package options](http://yihui.name/knitr/options) describe how caching works and the [cache examples](http://yihui.name/knitr/demo/cache/) provide additional details.

***
## [Continue to Inline Code](lesson-4.html)
