---
title: "Parameters"
output:
  html_document:
    toc_float: false
    includes:
      before_body: [include_header.html, include_lessons_nav.html]
      after_body: [include_lessons_nav_end.html, include_footer.html]
---

R Markdown documents can optionally include one or more parameters whose values can be set when you render the report. Parameters are useful when you want to re-render the same report with distinct values for various key inputs, for example:

* Running a report specific to a department or geographic region.

* Running a report that covers a specific period in time.

* Running multiple versions of a report for distinct sets of core assumptions.

## Declaring Parameters

Parameters are declared using the `params` field within the YAML section at the top of the document, for example:

```{r eval = FALSE}
---
title: My Document
output: html_document
params:
   data: "florida"
---
```

### Parameter Types

R Markdown can parse all of the standard R object types including character, integer, numeric, and logical, as well as R expressions. To pass an R expression, preface the parameter value with `!r` to indicate that the value is an R expression rather than a literal value:

```{r eval = FALSE}
---
title: My Document
output: html_document
params:
  start: !r as.Date("2015-01-01")
  snapshot: !r as.POSIXct("2015-01-01 12:30:00")
---
```

## Using Parameters in Code

Parameter values are made available within the knit environment as components of a read-only list named `params`. For example, the values of the above two parameters can be accessed in a chunk with the following R code:

```{r eval = FALSE}
params$start
params$snapshot
```


## Setting Parameters values for render

R Markdown will use the default parameter values listed in the YAML when you preview a [document that uses parameters](demos/4-params.Rmd) in RStudio (or otherwise call `render` with no arguments).

<!-- Screenshot of 4-params.Rmd being rendered.
![]()
-->

You can vary the parameter values for a specific report by providing a list of `params` values when you call `render`:

```{r eval = FALSE}
rmarkdown::render("4-params.Rmd", params = list(data = "hawaii"))
```

<!-- Screenshot of 4-params.Rmd being rendered with hawaii.
![]()
-->

Better yet, you can clikc the "Knit with Parameters" option in the RStudio IDE knit to launch a user-friendly interface for setting parameters.

<!-- Screenshot of knit with parameters.
![]()
-->

You can learn more about parameters at [Paramaterized Reports](http://rmarkdown.rstudio.com/developer_parameterized_reports.html).

***

## [Continue to Markdown Basics](lesson-7.html)
